---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rmarkdown.checks

<!-- badges: start -->
<!-- badges: end -->

With rmarkdown.checks you can validate input for parameterized Rmd files before trying to knit them.

## Why 

You are using parameters on your Rmarkdown documents but you are not using RStudio's  interactive user interface, so [these instructions](https://bookdown.org/yihui/rmarkdown/params-knit.html) don't apply. 

You  want to put constrains on the parameters supplied and have good reason wanting to test for them before trying to knit the report (e.g. because knitting happens on another machine).

## Installation

``` r
remotes::install_github("benjaminschwetz/rmarkdown.checks")
```

## Example

A `Rmd` document with the following header:

```yaml
---
title: "Parameter Checks"
author: "Benjamin Schwetz"
date: "5/31/2022"
output: html_document
params:
  fruit:
    value: Orange
    check: is.character
  other:
    value: some
---
```

can be inspected like so:

```{r example}
library(rmarkdown.checks)
check_parameters(
  rmarkdown_document = system.file("extdata", "sample.Rmd", package = "rmarkdown.checks"),
    params = list(fruit = "Apple"),
    allow_NULL = FALSE,
    allow_NA = FALSE,
    verbose = TRUE
)
```

## Background

Rmarkdown reports can be parameterised using the yaml header. The simple way is 

```yaml
---
title: My Document
output: html_document
params:
  fruit: Apple
---
```

Alternatively,  values can be declared using a `value` subitem like so:

```yaml
---
title: My Document
output: html_document
params:
  fruit: 
    value: Apple
---
```

Additional arbitrary subitems are also allowed. This is used for [customizing the interactive user interface](https://bookdown.org/yihui/rmarkdown/params-knit.html)


In this package, we use this feature to declare a `check` subitems to declare test functions:

## Writing `check` functions

### `check` functions must evaluate to `TRUE` or `FALSE`

The check function must take the `value` as first argument and always return either `TRUE` or `FALSE`. This makes it a bit different from normal tests and is also why I am trying to avoid the word test here.

There are two ways to write a check:

### Reference a function by name 

```yaml
---
title: My Document
output: html_document
params:
  fruit: 
    value: Apple
    check: is.character
---
```

### Declare a function

```yaml
---
title: My Document
output: html_document
params:
  fruit: 
    value: Apple
    check: \(x){is.character(x)}
---
```

Or the long way:

```yaml
---
title: My Document
output: html_document
params:
  fruit: 
    value: Apple
    check: function(x){is.character(x)}
---
```

## I need to run multiple checks or need to test for something more complicated!

I am trying to keep things intentionally simple so that the headers remain human readable.

For things that don't fit one line, I would declare the function in a package and run

```yaml
---
title: My Document
output: html_document
params:
  fruit: 
    value: Apple
    check: my_package::my_check
---
```
